Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula

### 2015-03-29 numerical.tex
major revisions

\subsection{Vertically non-isothermal disks}

>>>>

\subsection{Vertically non-isothermal disks} 
%We briefly consider vertically non-isothermal disks in which there is
%a disk surface at finite height. 
\cite{barker15} have argued that vertically isothermal disks are not
ideal for studing the VSI because $\zmax$ is a free
parameter in numerical calcluations. On the other hand, vertically
non-isothermal 
disks have a disk surface at finite height so there is a natural
choice for $\zmax$. This means that higher order modes
and surface modes are better defined in a vertically non-isothermal
disk. However, note that this issue is unimportant for low order modes 
such as the fundamental mode, which are not sensitive to boundary
conditions.

Here we briefly consider a vertically non-isothermal disk model with  
$(p,q,\epsilon)=(0,-1,0.05)$ and $\Gamma=1.0837$. This value of 
$\Gamma$ is chosen so that the standard vertical domain size $\zmax =
5H$ effectively corresponds to the disk surface ($\zmax\simeq
H_s$). The disk temperature decreases rapidly with height such that 
$c_s(\zmax)\simeq 10^{-4}c_s(0)$.

In Fig. \ref{compare_modes_vnoniso_kx10} we plot the mode diagram for
$\khat=10$ for several values of $\beta$. This plot is
qualitatively similar to Fig. \ref{compare_modes_cool_kx10}. As before
we find as $\beta$ is increased, higher order modes are rapidly
stabilized, eventually only the fundamental mode can operate before
the VSI is entirely stabilized.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
  0cm]{figures/compare_modes_Gam1.084_kx10.ps}
  \caption{Unstable modes in a vertically non-isothermal disk with
    $\khat=10$ and a range of thermal relaxation timescales.  
    \label{compare_modes_vnoniso_kx10}}
\end{figure}

We do not find surface modes for $\khat=10$. In the vertically
non-isothermal disk, surface modes develop for sufficiently large 
$\khat$ \citep{barker15}. We demonstrate them in
Fig. \ref{compare_modes_vnoniso_kx100} which show eigenvalues for
$\khat=100$. While growth rates generally decrease with increasing 
$\beta$, we find eigenvalues can overlap, making it difficult to 
distinguish between different modes. However, since there 
is a well-defined surface, we may seek the 
most unstable mode without the boundary-dependent issue that would
arise in a vertically isothermal disk.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
  0cm]{figures/compare_modes_Gam1.084_kx100.ps}
  \caption{Same as Fig. \ref{compare_modes_vnoniso_kx10} but with
    $\khat=100$. In this case surface modes appear and are readily identified for
    $\beta=10^{-3}$ and nearly
    parallel to the vertical axis
    \label{compare_modes_vnoniso_kx100}}
\end{figure}

Fig. \ref{compare_eigenvz_vnoniso_kx100}

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.73cm 0cm
  0cm]{figures/eigenvectorvz_vnoniso_kx100_beta0d001.ps}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.73cm 0cm
  0cm]{figures/eigenvectorvz_vnoniso_kx100_beta0d01.ps}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
  0cm]{figures/eigenvectorvz_vnoniso_kx100_beta0d1.ps}
  \caption{Vertical velocity eigenfunction of the most unstable mode
    with $\khat=100$ found in the vertically non-isothermal disk with
    $\beta = 10^{-3}$ (top), $\beta=10^{-2}$ (middle) and $\beta=0.1$
    (bottom). 
    \label{compare_eigenvz_vnoniso_kx100}}
\end{figure}
